On-the-fly Algorithms for GKAT Equivalences

### 2025-01-24 main.tex
making table less crammed
minor fix

%%%%% PACKAGES
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[margin=2.5cm]{geometry}

@@

This surprising fact lead to the development of Guarded Kleene Algebra with Tests (GKAT)~\cite{smolka_GuardedKleeneAlgebra_2020}, a (co)algebraic system that seeks to characterize this efficient fragment of KAT.
The determinism of GKAT gives it a different coalgebra structure from KAT, and allows for a GKAT expression to be directly converted to a linear-sized coalgebra, as opposed to the exponential sized coalgebra produced by KAT expressions~\cite{smolka_GuardedKleeneAlgebra_2020,kozen_CoalgebraicTheoryKleene_2017}.

@@

\begin{proofEnd}
    Because homomorphic image preserves principle sub-GKAT coalgebra (\Cref{thm:homo-img-preserve-principle-sub-coalg})
    \[h(⟨s⟩_S) = ⟨h(s)⟩_U;\]
    therefore for any state \(s' ∈ S\):
    \[s' ∈ ⟨s⟩_S ⟺ h(s') ∈ h(⟨s⟩_S) ⟺ h(s') ∈ ⟨h(s)⟩_U.\]
    And because \(s'\) is accepting if and only if \(h(s')\) accepting by definition of homomorphism; then \(⟨s⟩_S\) contains an accepting state if and only if \(⟨h(s)⟩_U\) contains an accepting state. 
    Therefore, \(s\) is live in \(S\) if and only if \(h(s)\) is live in \(U\).
\end{proofEnd}

@@

\begin{theoremEnd}{corollary}[bisimulation preserves liveness]\label{thm:bisim-preserve-liveness}
    If there exists a bisimulation \(∼\) between GKAT coalgebra \(S\) and \(U\) s.t. \(s ∼ u\) for some states \(s ∈ S\) and \(u ∈ U\), then \(s\) and \(u\) has to be either both accepting, both live or both dead.
\end{theoremEnd}

@@

\begin{table*}
    \centering
    \begin{tabular}{c||c|c|l|l}
        Exp & \(S\) & \(s^*\)  
        & \(ϵ̂(s)\) & \(δ̂(s)\) \\\hline
        \(b\) & \(\{s^*\}\) 
        & \(s^*\)
        & \(\{b\}\) & \(∅\) \\  
        \(p\) & \(\{s^*, s₁\}\) 
        & \(s^*\)
        & \(\begin{cases}
           ∅ & s = s^* \\  
           \{1\} & s = s₁ 
        \end{cases}\) 
        & \(\begin{cases}
            \{(1, s₁, 0)\} & s = s^* \\  
            ∅ & s = s₁
        \end{cases}\)\\  
        \(e₁ +_b e₂\) & \(\{s^*\} + S₁ + S₂\) &
        \(s^*\) &
        \(\begin{cases}
            ⟨\{b\}| ϵ̂₁(s₁^*) ∪ ⟨\{b\}| ϵ̂₂(s₂^*) & s = s^* \\
            ϵ̂₁(s) & s ∈ S₁\\
            ϵ̂₂(s) & s ∈ S₂\\
        \end{cases}\) & 
        \(\begin{cases}
            ⟨\{b\}| δ̂₁(s₁^*) + ⟨\{b\}| δ̂₂(s₂^*) & s = s^* \\
            δ̂₁(s) & s ∈ S₁\\
            δ̂₂(s) & s ∈ S₂\\
        \end{cases}\) \\  
        \(e₁ ; e₂\) & \(S₁ + S₂\) & 
        \(s₁^*\) & 
        \(\begin{cases}
            ⟨ϵ̂₁(s)| ϵ̂₂(s₂^*)& s ∈ S₁ \\  
            ϵ̂₂(s) & s ∈ S₂
        \end{cases}\)& 
        \(\begin{cases}
            δ̂₁(s) + ⟨ϵ̂(s)| δ̂₂(s₂^*) & s ∈ S₁ \\  
            δ̂2(s) & s ∈ S₂
        \end{cases}\) \\  
        \(e₁^{(b)}\) & \(\{s^*\} + S₁\) & 
        \(s^*\) &
        \(\begin{cases}
            \{\overline{b}\} & s = s^*\\
            ⟨\{\overline{b}\}| ϵ̂1(s) & s ∈ S₁
        \end{cases}\) & 
        \(\begin{cases}
            ⟨\{b\}| δ̂₁(s₁^*) & s = s^* \\
            δ₁(s) ∪ ⟨\{b\}|⟨ϵ̂1(s)| δ̂₁(s₁^*) & s ∈ S₁
        \end{cases}\)
    \end{tabular}
    \caption{Symbolic Thompson's Construction}\label{tab:symb-Thompson-construction}
\end{table*}

>>>>

%%%%% PACKAGES
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[margin=2.5cm]{geometry}
\usepackage{tabularx}

@@

This surprising fact leads to the development of Guarded Kleene Algebra with Tests (GKAT)~\cite{smolka_GuardedKleeneAlgebra_2020}, a (co)algebraic system that seeks to characterize this efficient fragment of KAT.
The determinism of GKAT gives it a different coalgebra structure from KAT, and allows for a GKAT expression to be directly converted to a linear-sized coalgebra, as opposed to the exponential sized coalgebra produced by KAT expressions~\cite{smolka_GuardedKleeneAlgebra_2020,kozen_CoalgebraicTheoryKleene_2017}.

@@

\begin{proofEnd}
    Because homomorphic image preserves principle sub-GKAT coalgebra (\Cref{thm:homo-img-preserve-principle-sub-coalg}) \(h(⟨s⟩_S) = ⟨h(s)⟩_U;\) therefore for any state \(s' ∈ S\):
    \[s' ∈ ⟨s⟩_S ⟺ h(s') ∈ h(⟨s⟩_S) ⟺ h(s') ∈ ⟨h(s)⟩_U.\]
    And because \(s'\) is accepting if and only if \(h(s')\) accepting by definition of homomorphism; then \(⟨s⟩_S\) contains an accepting state if and only if \(⟨h(s)⟩_U\) contains an accepting state. 
    Therefore, \(s\) is live in \(S\) if and only if \(h(s)\) is live in \(U\).
\end{proofEnd}

@@

\begin{theoremEnd}{corollary}[bisimulation preserves liveness]\label{thm:bisim-preserve-liveness}
    For two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ U\), with a bisimulation \({∼} ⊆ S × U\) s.t. \(s ∼ u\), then \(s\), \(u\), and \((s, u)\) has the same liveness, i.e. needs to be all accepting, all live, or all dead, in \(S\), \(U\), and \(∼\) respectively.
\end{theoremEnd}

@@

\begin{table*}
    \centering
    \setlength{\extrarowheight}{3px}
    \begin{tabular}{c||c|c|l|l}
        Exp & \(S\) & \(s^*\)  
        & \(ϵ̂(s)\) & \(δ̂(s)\) \\\hline
        \(b\) & \(\{s^*\}\) 
        & \(s^*\)
        & \(\{b\}\) & \(∅\) \\  
        \(p\) & \(\{s^*, s₁\}\) 
        & \(s^*\)
        & \(\begin{cases}
           ∅ & s = s^* \\  
           \{1\} & s = s₁ 
        \end{cases}\) 
        & \(\begin{cases}
            \{(1, s₁, 0)\} & s = s^* \\  
            ∅ & s = s₁
        \end{cases}\)\\  
        \(e₁ +_b e₂\) & \(\{s^*\} + S₁ + S₂\) &
        \(s^*\) &
        \(\begin{cases}
            ⟨\{b\}| ϵ̂₁(s₁^*) ∪ ⟨\{b\}| ϵ̂₂(s₂^*) & s = s^* \\
            ϵ̂₁(s) & s ∈ S₁\\
            ϵ̂₂(s) & s ∈ S₂\\
        \end{cases}\) & 
        \(\begin{cases}
            ⟨\{b\}| δ̂₁(s₁^*) + ⟨\{b\}| δ̂₂(s₂^*) & s = s^* \\
            δ̂₁(s) & s ∈ S₁\\
            δ̂₂(s) & s ∈ S₂\\
        \end{cases}\) \\  
        \(e₁ ; e₂\) & \(S₁ + S₂\) & 
        \(s₁^*\) & 
        \(\begin{cases}
            ⟨ϵ̂₁(s)| ϵ̂₂(s₂^*)& s ∈ S₁ \\  
            ϵ̂₂(s) & s ∈ S₂
        \end{cases}\)& 
        \(\begin{cases}
            δ̂₁(s) + ⟨ϵ̂(s)| δ̂₂(s₂^*) & s ∈ S₁ \\  
            δ̂2(s) & s ∈ S₂
        \end{cases}\) \\  
        \(e₁^{(b)}\) & \(\{s^*\} + S₁\) & 
        \(s^*\) &
        \(\begin{cases}
            \{\overline{b}\} & s = s^*\\
            ⟨\{\overline{b}\}| ϵ̂1(s) & s ∈ S₁
        \end{cases}\) & 
        \(\begin{cases}
            ⟨\{b\}| δ̂₁(s₁^*) & s = s^* \\
            δ₁(s) ∪ ⟨\{b\}|⟨ϵ̂1(s)| δ̂₁(s₁^*) & s ∈ S₁
        \end{cases}\)
    \end{tabular}
    \caption{Symbolic Thompson's Construction}\label{tab:symb-Thompson-construction}
\end{table*}
